---
title: Joining Data Frames with sf Objects
---

### Import CSV

```{r}
library(tidyverse)

city_councilors <-
  read_csv("data/portland_city_councilors.csv")
```

```{r}
city_councilors
```

### Import `sf` Object


```{r}
library(sf)

city_council_districts <-
  read_sf("data/portland_city_council_districts.geojson")

city_council_districts
```

---

```{r}
library(mapview)

city_council_districts |>
  mapview()
```

### Joing CSV and `sf` Object {.inverse}

#### `left_join()`

```{r}
left_join(
  city_councilors,
  city_council_districts,
  join_by(district)
)
```

---

```{r}
left_join(
  city_councilors,
  city_council_districts,
  join_by(district)
) |>
  st_as_sf()
```

---

```{r}
left_join(
  city_councilors,
  city_council_districts,
  join_by(district)
) |>
  st_as_sf() |>
  mapview()
```

#### `right_join()`

```{r}
right_join(
  city_councilors,
  city_council_districts,
  join_by(district)
) |>
  st_as_sf()
```


### Your Turn {.your-turn}

```{r}
#| eval: false
#| echo: false

library(tidyverse)
library(sf)
library(janitor)

read_csv(
  "https://hub.arcgis.com/api/v3/datasets/437a2765164d4f01b23d463e0e2135a1_54/downloads/data?format=csv&spatialRefId=3857&where=1%3D1"
) |>
  clean_names()

traffic_signals <-
  read_sf(
    "https://hub.arcgis.com/api/v3/datasets/437a2765164d4f01b23d463e0e2135a1_54/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1"
  ) |>
  clean_names()

traffic_signals |>
  st_drop_geometry() |>
  select(objectid, power_supplier) |>
  write_csv("data/portland_traffic_signals_power_supplier.csv")

traffic_signals |>
  select(objectid) |>
  write_sf("data/portland_traffic_signals.geojson")
```

- Import geospatial data on Portland traffic signals using the following code:

. . .

```{r}
#| eval: false
library(sf)

portland_traffic_signals <-
  read_sf(
    "https://raw.githubusercontent.com/rfortherestofus/mapping-with-r-v2/refs/heads/main/data/portland_traffic_signals.geojson"
  )
```

### Your Turn {.your-turn}

- Import data on the power supplier of Portland traffic signals in a CSV file using the following code:

. . .

```{r}
#| eval: false
library(tidyverse)

portland_traffic_signals_power_supplier <-
  read_csv(
    "https://raw.githubusercontent.com/rfortherestofus/mapping-with-r-v2/refs/heads/main/data/portland_traffic_signals_power_supplier.csv"
  )
```

### Your Turn {.your-turn}

- Join the `portland_traffic_signals` and `portland_traffic_signals_power_supplier` objects

- Try making a map with different power suppliers in different colors to ensure that your join worked
