---
title: Adding Text to Maps
---

```{r}
#| output: false
#| echo: false
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)

speak_language_other_than_english <-
  get_acs(
    geography = "county",
    variable = "S1601_C01_003",
    summary_var = "S1601_C01_001",
    geometry = TRUE
  ) |>
    clean_names() |>
    mutate(pct = estimate / summary_est) |>
    select(name, pct)

languages_map <-
  speak_language_other_than_english |>
    shift_geometry() |>
    ggplot() +
    geom_sf(
      aes(
        fill = pct,
        color = pct
      )
    ) +
    guides(
      fill = guide_colorsteps(show.limits = TRUE),
      color = guide_colorsteps(show.limits = TRUE),
    ) +
    scale_fill_viridis_c(
      limits = c(0, 1),
      labels = percent_format()
    ) +
    scale_color_viridis_c(
      limits = c(0, 1),
      labels = percent_format()
    ) +
    labs(
      color = NULL,
      fill = NULL,
      title = "Percentage of People Who Speak a Language\nOther than English in Each County in the United States"
    ) +
    theme_void() +
    theme(
      legend.position = "top",
      legend.key.width = unit(2, "cm"),
      legend.key.height = unit(0.5, "cm"),
      plot.title = element_text(
        hjust = 0.5,
        face = "bold",
        size = 16,
        margin = margin(
          b = 0.25,
          unit = "cm"
        )
      )
    )
```



### Add Text to Map

```{r}
#| output: false
oregon_counties <-
  counties(state = "OR") |>
    clean_names()
```

```{r}
oregon_counties
```

---

```{r}
oregon_counties |>
  ggplot() +
  geom_sf() +
  geom_sf_text(aes(label = name))
```

### Map from Last Lesson 

---

```{r}
languages_map
```

::: {.notes}
Show this map: https://show.rfor.us/XlwwHHkF ([source](https://links.mimestream.com/g/david@rfortherestofus.com/t/194b89b5675c462b))
:::

### `geom_sf_text()`

```{r}
speak_language_other_than_english
```

---

```{r}
#| output-location: slide
languages_map +
  geom_sf_text(aes(label = name))
```

### Only Show Some Text

---

```{r}
library(sf)

top_counties <-
  speak_language_other_than_english |>
    separate_wider_delim(name, delim = ", ", names = c("county", "state")) |>
    filter(state != "Puerto Rico") |>
    slice_max(order_by = pct, n = 5) |>
    mutate(map_label = str_glue("{county} ({percent(pct, accuracy = 1)})")) |>
    st_as_sf()
```

---

```{r}
top_counties
```

---

```{r}
#| output-location: slide
languages_map +
  geom_sf_text(
    data = top_counties,
    aes(label = map_label)
  )
```

### Use {ggrepel} to Avoid Overlapping Text

---

```{r}
#| output-location: slide
#| code-line-numbers: "4,10"
library(ggrepel)

languages_map +
  geom_text_repel(
    data = top_counties,
    aes(
      label = map_label,
      geometry = geometry
    ),
    stat = "sf_coordinates"
  )
```

### Additional Tweaks to Improve {ggrepel}

```{r}
#| output-location: slide
#| code-line-numbers: "9-16"
languages_map +
  geom_text_repel(
    data = top_counties,
    aes(
      label = map_label,
      geometry = geometry
    ),
    stat = "sf_coordinates",
    seed = 1234,
    box.padding = unit(10, "pt"),
    segment.color = "gray30",
    segment.curvature = -0.1,
    arrow = arrow(length = unit(2, "pt"), type = "closed"),
    color = "white",
    bg.color = "grey30",
    bg.r = 0.15
  )
```

### Your Turn {.your-turn}

Improve the map you made in the last lesson on the number of refugees from each country in the world by:

1. Calculating the three top countries refugee-sending countries
1. Adding text to the map to highlight these countries

### Learn More

- Link to highlighting with shadows blog post
