---
title: Choosing a Projection
---

### All Projections are Wrong

::: {.notes}
Show clip from this video: https://www.youtube.com/watch?v=kIID5FDi2JQ
:::

### Show a Bunch of Projections

::: {.notes}
https://map-projections.net/single-view/canters-francula

https://www.reddit.com/r/dataisbeautiful/comments/9nkhkz/animating_the_mercator_projection_to_the_true/

https://mrgris.com/projects/merc-extreme/

https://en.wikipedia.org/wiki/Cahill%E2%80%93Keyes_projection#/media/File:Cahill-Keyes_projection.jpg

https://bsky.app/profile/pokateo.bsky.social/post/3lhccfxdnsk2v
:::

---

```{r}
library(tidyverse)
library(rnaturalearth)

all_countries <-
  ne_countries() |>
    select(sovereignt)
```

. . .

```{r}
all_countries
```

---

```{r}
#| output-location: slide

library(sf)

all_countries |>
  ggplot() +
  geom_sf()
```


#### Robinson Projection

```{r}
#| code-line-numbers: "2"
#| output-location: slide
all_countries |>
  st_transform(
    "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
  ) |>
  ggplot() +
  geom_sf()
```

### Gall-Peters Projection

```{r}
#| code-line-numbers: "2"
#| output-location: slide
all_countries |>
  st_transform("+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
  ) |>
  ggplot() +
  geom_sf()
```


#### Mollweide Projection

```{r}
#| code-line-numbers: "2"
#| output-location: slide
all_countries |>
  st_transform("+proj=moll +datum=WGS84 +units=m") |>
  ggplot() +
  geom_sf()
```


#### Lambert Azimuthal Equal-Area Projection

```{r}
#| code-line-numbers: "2"
#| output-location: slide
all_countries |>
  st_transform("+proj=laea +x_0=0 +y_0=0 +lon_0=0 +lat_0=0") |>
  ggplot() +
  geom_sf()
```

### {crsuggest}

```{r}
#| output-location: slide
library(crsuggest)

africa <-
  ne_countries(continent = "Africa") |>
    select(sovereignt)

africa |>
  ggplot() +
  geom_sf()
```

---

```{r}
africa |>
  suggest_top_crs()
```

. . .

```{r}
africa_crs <-
  africa |>
    suggest_top_crs()
```

. . .

```{r}
#| code-line-numbers: "2"
#| output-location: slide

africa |>
  st_transform(africa_crs) |>
  ggplot() +
  geom_sf()
```

---

:::: {.columns}

::: {.column width="50%"}
```{r}
#| echo: false
africa |>
  ggplot() +
  geom_sf()
``
`
:::

::: {.column width="50%"}

```{r}
#| echo: false
africa |>
  st_transform(africa_crs) |>
  ggplot() +
  geom_sf()
```

:::

::::

---

```{r}
#| echo: false
#| output: false
library(tigris)

continental_us_states <-
  states() |>
  select(NAME) |>
  filter(NAME %in% state.name) |>
  filter(NAME != "Alaska") |>
  filter(NAME != "Hawaii")
```

:::: {.columns}

::: {.column width="50%"}

```{r}
#| echo: false
continental_us_states |>
  ggplot() +
  geom_sf()
```

:::

::: {.column width="50%"}

```{r}
#| echo: false
us_crs <-
  continental_us_states |>
  suggest_top_crs()

continental_us_states |>
  st_transform(us_crs) |>
  ggplot() +
  geom_sf()
`
``
:::

::::

### Start with WGS84

> When selecting geographic CRSs, the answer is often WGS84. ... WGS84 is the most common CRS in the world, so it is worth knowing its EPSG code: 4326. 

### Your Turn {.your-turn}

- Improve the map you made in the last lesson on the number of refugees from each country in the world by using the `crssuggest` package to find the best projection for the data.

- Once you've done this, redo your map with this projection.

### Learn More

https://www.youtube.com/watch?v=kIID5FDi2JQ

https://r.geocompx.org/reproj-geo-data.html
