---
title: Making Interactive Maps with {mapgl}
eval: false
---

::: {.notes}
Can't render the code in these slides because can't render Mapbox in Quarto
See https://github.com/walkerke/mapgl/issues/3
:::

```{r}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)
library(sf)
library(mapgl)
```

### {mapgl} {.inverse}

::: {.notes}
Show mapbox website
:::

### Get Mapbox Access Token

::: {.notes}
Show creating account and access token
Add access token to Renviron
:::

### Basic Map {.inverse}

#### Styles {.inverse}

::: {.notes}
Mention how can't render Mapbox maps in Quarto at the moment (https://github.com/walkerke/mapgl/issues/3) but cant use maplibre
:::

---

```{r}
mapboxgl(style = mapbox_style("light"))
```

---

```{r}
mapboxgl(style = mapbox_style("dark"))
```

---

```{r}
mapboxgl(style = mapbox_style("outdoors"))
```

---

```{r}
maplibre()
```

---

```{r}
maplibre(style = carto_style("positron"))
```

---
```{r}
maplibre(style = carto_style("dark-matter"))
```

#### Adjust Center and Zoom

```{r}
mapboxgl(
  mapbox_style("light"),
  center = c(-43.23412, -22.91370)
)
```

---

```{r}
mapboxgl(
  mapbox_style("light"),
  center = c(-43.23412, -22.91370),
  zoom = 5
)
```

#### Add Layers

```{r}
speak_language_other_than_english_wgs84_with_labels <-
  read_sf("data/speak_language_other_than_english_wgs84_with_labels.geojson")
```

---

```{r}
mapboxgl(
  mapbox_style("light"),
  center = c(-100.29957845401096, 38.93484429077689),
  zoom = 2
) |>
  add_fill_layer(
    id = "speak_language_other_than_english_wgs84_with_labels",
    source = speak_language_other_than_english_wgs84_with_labels
  )
```

---

```{r}
mapboxgl(
  mapbox_style("light"),
  center = c(-100.29957845401096, 38.93484429077689),
  zoom = 2
) |>
  add_fill_layer(
    id = "speak_language_other_than_english_wgs84_with_labels",
    source = speak_language_other_than_english_wgs84_with_labels,
    fill_color = interpolate(
      column = "pct",
      values = c(0, 1),
      stops = c("lightblue", "darkblue"),
      na_color = "lightgrey"
    ),
    fill_opacity = 1
  )
```

```{r}
mapboxgl(
  mapbox_style("light"),
  center = c(-100.29957845401096, 38.93484429077689),
  zoom = 2
) |>
  add_fill_layer(
    id = "speak_language_other_than_english_wgs84_with_labels",
    source = speak_language_other_than_english_wgs84_with_labels,
    fill_color = interpolate(
      column = "pct",
      values = c(0, 1),
      stops = c("lightblue", "darkblue"),
      na_color = "lightgrey"
    ),
    popup = "text_label",
    fill_opacity = 1
  )
```

### Other Layers {.inverse}

::: {.notes}
Show layers on https://walker-data.com/mapgl/articles/layers-overview.html
:::

### Other Things You Can Do with {mapgl}

```{r}
mapboxgl() |>
  fly_to(
    center = c(2.1228412549747344, 41.3812461533571),
    zoom = 16,
    pitch = 25
  )
```

### Your Turn {.your-turn}

- Turn any of the interactive refugee maps you have made into an interactive map with {mapgl}