---
title: Heatmaps
---

::: {.notes}
Example: https://docs.mapbox.com/mapbox-gl-js/exafmple/heatmap-layer/
:::

### Load Data

```{r}
library(sf)
library(tidyverse)
library(janitor)

improved_corners <-
  read_sf("data/improved_corners.geojson")

improved_corners
```


```{r}
#| eval: false
#| echo: false
# From https://gis-pdx.opendata.arcgis.com/datasets/city-boundaries/about

# read_sf(
#   "https://hub.arcgis.com/api/v3/datasets/951488174bfe4275bbbd04421d7820f5_10/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1"
# ) |>
#   clean_names() |>
#   filter(cityname == "Portland") |>
#   st_make_valid() |>
#   write_sf("data/portland_boundaries.geojson")
```

```{r}
portland_boundaries <-
  read_sf("data/portland_boundaries.geojson")
```

### Make Grid

```{r}
portland_boundaries |>
  st_make_grid()
```

```{r}
portland_boundaries |>
  st_make_grid() |>
  ggplot() +
  geom_sf()
```

```{r}
portland_boundaries |>
  st_make_grid(n = c(20, 20))
```

```{r}
portland_boundaries |>
  st_make_grid(n = c(20, 20)) |>
  ggplot() +
  geom_sf()
```

```{r}
portland_grid <-
  portland_boundaries |>
  st_make_grid(n = c(50, 50))
```

```{r}
ggplot() +
  geom_sf(data = portland_boundaries) +
  geom_sf(data = portland_grid, alpha = 0.5) +
  theme_void()
```

```{r}
ggplot() +
  geom_sf(data = portland_boundaries) +
  geom_sf(data = portland_grid, alpha = 0.5)
```

### Clip Map to Portland Boundaries
```{r}
portland_grid_map <-
  st_intersection(portland_boundaries, portland_grid) |>
  mutate(grid_id = row_number()) |>
  select(grid_id)
```

```{r}
ggplot() +
  geom_sf(data = portland_grid_map) +
  theme_void()
```

### Join Corners Data with Grid Map

```{r}
improved_corners_grid <-
  st_join(
    portland_grid_map,
    improved_corners
  )

improved_corners_grid
```

### Calculate Percentage of Unimproved Corners in Each Square 
```{r}
improved_corners_grid_pct <-
  improved_corners_grid |>
  st_drop_geometry() |>
  count(grid_id, ramp_style) |>
  complete(grid_id, ramp_style) |>
  group_by(grid_id) |>
  mutate(pct = n / sum(n, na.rm = TRUE)) |>
  ungroup() |>
  select(-n) |>
  pivot_wider(
    id_cols = grid_id,
    names_from = ramp_style,
    values_from = pct
  ) |>
  mutate(
    pct = case_when(
      is.na(Unimproved) & Improved == 1 ~ 1,
      .default = Unimproved
    )
  ) |>
  select(grid_id, pct) |>
  left_join(portland_grid_map, join_by(grid_id)) |>
  st_as_sf()
```

::: {.notes}
This code is quite complicated so show it piece by piece
:::

### Make Map
```{r}
library(scales)

ggplot() +
  geom_sf(data = portland_boundaries) +
  geom_sf(
    data = improved_corners_grid_pct,
    aes(fill = pct),
    color = "white"
  ) +
  labs(fill = NULL, title = "Percentage of unimproved corners in Portland") +
  scale_fill_viridis_c(
    option = "E",
    na.value = "gray90",
    limits = c(0, 1),
    labels = percent_format()
  ) +
  theme_void() +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      margin = margin(b = 10, unit = "pt")
    ),
    legend.key.width = unit(1.5, "cm"),
    legend.key.height = unit(0.5, "cm"),
    legend.position = "top"
  )
```

### Your Turn {.your-turn}

- Create a heatmap that shows where in Portland there are the most traffic signals (hint: it should be in the Southwest sextant, which includes downtown)

### Learn More

https://www.andrewheiss.com/blog/2023/07/28/gradient-map-fills-r-sf/