---
title: Choosing a Projection
---

### All Projections are Wrong

::: {.notes}
Show clip from this video: https://www.youtube.com/watch?v=kIID5FDi2JQ
:::

### Show a Bunch of Projections

::: {.notes}
https://map-projections.net/single-view/canters-francula

https://www.reddit.com/r/dataisbeautiful/comments/9nkhkz/animating_the_mercator_projection_to_the_true/

https://bsky.app/profile/locations.bsky.social/post/3lgpvzjcmek2k
:::

```{r}
library(rnaturalearth)

all_countries <-
  ne_countries()
```

```{r}
library(sf)
library(tidyverse)
```

```{r}
all_countries |>
  st_transform("+proj=laea +x_0=0 +y_0=0 +lon_0=0 +lat_0=0") |>
  ggplot() +
  geom_sf()
```

```{r}
all_countries |>
  st_transform(4267) |>
  ggplot() +
  geom_sf()
```

```{r}
all_countries |>
  crsuggest::suggest_top_crs()

all_countries |>
  st_transform(3395) |>
  ggplot() +
  geom_sf()
```

### Start with WGS84

> When selecting geographic CRSs, the answer is often WGS84. It is used not only for web mapping, but also because GPS datasets and thousands of raster and vector datasets are provided in this CRS by default. WGS84 is the most common CRS in the world, so it is worth knowing its EPSG code: 4326.32 This ‘magic number’ can be used to convert objects with unusual projected CRSs into something that is widely understood.

### Learn More

https://www.youtube.com/watch?v=kIID5FDi2JQ

https://r.geocompx.org/reproj-geo-data.html?q=project#which-crs