---
title: Accessing Geospatial Data Through Packages
execute: 
  freeze: true
  eval: false
---

```{r}
#| echo: false
library(tidycensus)

options(tigris_use_cache = TRUE)
```

### Overview

- Most of the time, I don't access raw geospatial data
- Instead, I use packages to access the data

### Global Data

#### Boundaries are Political

#### Packages

- {rnaturalearth}
- {rgeoboundaries}

::: {.notes}
Talk about how each package handles disputed territories
:::

#### {rnaturalearth}

---

```{r}
library(rnaturalearth)
```

```{r}
all_countries <-
  ne_countries()
```

---

```{r}
library(mapview)

all_countries |>
  mapview()
```

---

```{r}
africa <-
  ne_countries(continent = "Africa")
```

```{r}
africa |>
  mapview()
```

---

```{r}
ukraine <-
  ne_countries(country = "Ukraine")
```

```{r}
ukraine |>
  mapview()
```

#### {rgeoboundaries}

::: {.notes}
Show https://www.geoboundaries.org/

https://github.com/wmgeolab/rgeoboundaries
:::

---

```{r}
library(rgeoboundaries)
```

```{r}
all_countries_v2 <-
  geoboundaries(adm_lvl = 0)
```

```{r}
all_countries_v2 |>
  mapview()
```

---

```{r}
ukraine_v2 <-
  geoboundaries(country = "Ukraine")
```

```{r}
ukraine_v2 |>
  mapview()
```

---

```{r}
geoboundaries(
  country = "Ukraine",
  adm_lvl = 1
) |>
  mapview()
```

---

```{r}
geoboundaries(
  country = "Ukraine",
  adm_lvl = 2
) |>
  mapview()
```

::: {.notes}
Can go down to 5 levels; see https://github.com/wmgeolab/rgeoboundaries/blob/master/R/geoboundaries.R#L57C1-L62C87
:::


#### Other Packages

https://r.geocompx.org/read-write#geographic-data-packages

### US Data

- {tigris}
- {tidycensus}

#### {tigris}

```{r}
library(tigris)
```

```{r}
us_states <-
  states()
```

```{r}
us_states
```

---

```{r}
us_states |>
  mapview()
```

---

```{r}
oregon_counties <-
  counties(state = "OR")
```

```{r}
oregon_counties
```

---

```{r}
oregon_counties |>
  mapview()
```

#### {tidycensus}

- [Create an API key](https://api.census.gov/data/key_signup.html) 
- Register API key

. . .

```{r}
#| eval: false
library(tidycensus)

census_api_key("<insert your API key here>", install = TRUE)
```

#### Identify Variables

```{r}
acs_variables <-
  load_variables(
    year = 2020,
    dataset = "acs5",
    cache = TRUE
  )
```

```{r}
library(tidyverse)

median_household_income_variable <-
  acs_variables |>
  filter(str_detect(label, "Median household income")) |>
  filter(geography == "county") |>
  pull(name)
```

---

```{r}
median_household_income <-
  get_acs(
    geography = "county",
    state = "OR",
    variable = median_household_income_variable
  )
```

```{r}
median_household_income
```

---

```{r}
median_household_income_sf <-
  get_acs(
    geography = "county",
    state = "OR",
    variable = median_household_income_variable,
    geometry = TRUE
  )
```

```{r}
median_household_income_sf
```

---

```{r}
median_household_income_sf |>
  mapview(zcol = "estimate")
```

### Your Turn {.your-turn}

- Use the {rnaturalearth} or {rgeoboundaries} package to download data for all countries in South America
- Use the {mapview} package to make a map in order to ensure the data was downloaded correctly
- If you are in the US (or work with US data), try out downloading data using {tidycensus}

#### Learn More

https://wiki.openstreetmap.org/wiki/Disputed_territories
