---
title: Flow Maps
---


### Load Packages

```{r}
library(tidyverse)
library(tidycensus)
library(tigris)
library(janitor)
library(sf)
```

::: {.notes}
Things covered:

- Centroids
- Turning sf objects into non-sf objects and mapping them


https://bsky.app/profile/infobeautiful.bsky.social/post/3lhvuokmc5s2q
:::

### Import Data

```{r}
inflow <-
  get_flows(
    geography = "county",
    state = "OR",
    county = "Deschutes",
    geometry = TRUE
  ) |>
  clean_names() |>
  filter(variable == "MOVEDIN") |>
  slice_max(order_by = estimate, n = 10) |>
  drop_na(geoid2) # To get rid of Asia

inflow
```

---

```{r}
inflow_origins <-
  inflow |>
  st_drop_geometry() |>
  st_set_geometry("centroid2") |>
  select(geoid2, full2_name, estimate)

inflow_origins
```

---

```{r}
inflow_origins_centroid <-
  inflow |>
  select(centroid1, centroid2) |>
  st_drop_geometry() |>
  st_set_geometry("centroid2") |>
  st_coordinates(centroid2) |>
  as_tibble() |>
  set_names("origin_x", "origin_y")

inflow_origins_centroid
```

---

```{r}
inflow_destination_centroid <-
  inflow |>
  select(geoid1, full1_name, estimate) |>
  st_coordinates() |>
  as_tibble() |>
  set_names("destination_x", "destination_y")

inflow_destination_centroid
```

---

```{r}
inflow_x_y <-
  inflow |>
  select(estimate) |>
  st_drop_geometry() |>
  bind_cols(inflow_origins_centroid) |>
  bind_cols(inflow_destination_centroid)

inflow_x_y
```

---

```{r}
origin_states <-
  inflow_origins |>
  st_drop_geometry() |>
  separate_wider_delim(
    full2_name,
    delim = ", ",
    names = c("county", "state")
  ) |>
  distinct(state) |>
  pull(state)

origin_states
```

---

```{r}
origin_states_sf <-
  states(progress_bar = FALSE) |>
  clean_names() |>
  select(name) |>
  filter(name %in% origin_states)

origin_states_sf
```

### Make Map

```{r}
#| output-location: slide
ggplot() +
  geom_sf(data = origin_states_sf, fill = "transparent") +
  geom_curve(
    data = inflow_x_y,
    aes(
      x = origin_x,
      xend = destination_x,
      y = origin_y,
      yend = destination_y
    ),
    color = "orange"
  ) +
  geom_sf(
    data = inflow,
    shape = 21,
    size = 3,
    fill = "white"
  ) +
  geom_sf(
    data = inflow_origins
  ) +
  scale_size_continuous(range = c(0.5, 2)) +
  scale_linewidth_continuous(range = c(0.5, 2)) +
  theme_void()
```

### Your Turn {.your-turn}

Your job is to create a flow map showing the top 5 countries of origin for asylum seekers in the United States in 2023. 

. . .

Download data on the top 5 countries of origin for asylum seekers:

. . .


```{r}
#| eval: false
library(sf)

asylum_seekers <-
  read_sf(
    "https://raw.githubusercontent.com/rfortherestofus/mapping-with-r-v2/refs/heads/main/data/asylum_seekers.geojson"
  )
```

### Your Turn {.your-turn}

Import geospatial data on all countries:


```{r}
countries <-
  read_sf(
    "https://raw.githubusercontent.com/rfortherestofus/mapping-with-r-v2/refs/heads/main/data/countries.geojson"
  )
```

### Your Turn {.your-turn}

Adapt the sample code from the lesson to make a map that shows the flows of asylum seekers from the top 5 countries to the United States. You'll need to:

1. Use the `st_centroid()` function to turn the `asylum_seekers` object into geospatial data with single points (save it as `asylum_seekers_centroid`).

1. Use the `st_coordinates()` function to turn `asylum_seekers_centroid` into an object with x and y variables for the coordinates (save it as `asylum_seekers_centroid_x_y`).

1. Create a `united_states_centroid` object by filtering the `countries` object to just include the United States and then use `st_centroid()` to find its center.

1. Use the `st_coordinates()` function to turn `united_states_centroid` into an object with x and y variables for the coordinates (save it as `united_states_centroid_x_y`).

### Your Turn {.your-turn}

1. Use `bind_cols()` to bring together the `asylum_seekers_centroid_x_y` and `united_states_centroid_x_y` objects into a single object (save it as `asylum_seekers_flow`).

1. Make your map, using `geom_sf` to add all countries as a background layer before using `geom_curve()` to show the flow of asylum seekers from the top 5 countries.
